---
title: "Practical Machine Learning Prediction Assigment"
author: "Edana Merchan"
date: "January 22, 2015"
output: html_document
---
# Practical Machine Learning Prediction Assigment

## Summary

The assigment for this class uses data collected by devices such as Jawbone Up, Nike FuelBand, and Fitbit about personal activity. This measurements are used to improve users health, to find patterns in their behavior. In this project, we are going to use the data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

The goal of this project is to predict the way in which they did the exercise. Using the "classe" variable in the training set. At the end the analysis should use the prediction model to predict 20 different test cases. 

## Data collection 

The training data for this project was collecting from here: 

<https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv>

The test data was collecting from here: 

<https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv>

The data for this project come from this source: <http://groupware.les.inf.puc-rio.br/har>. 

```{r}
#Load the training and testing data treating empty values as NA
training <- read.csv("/Users/edana/Data_Fitbit/pml-training.csv", na.strings=c("NA",""))
testing <- read.csv("/Users/edana/Data_Fitbit/pml-testing.csv", na.strings=c("NA",""))
```

## Analysis
The main variable that we are interersted in is Class which gives the classification according how was performed the excersise:
- **A:** Exactly according to the specification.
- **B:** Throwing the elbows to the front.
- **C:** Lifting the dumbbell only halfway. 
- **D:** Lowering the dumbbell only halfway. 
- **E:** Throwing the hips to the front.

The idea is with the data collected predict if the activity is performed correctly.

There are many columns that are empty that have to ve cleaned to be able to have the ones that really represent the features in our problem.

```{r}
#Clean data from NA columns
features <- training[,colSums(is.na(training))<1]
testing <- testing[,colSums(is.na(testing))<1]

#The first 2 columns don't have useful information
features <- features[,8:60]
testing1 <- testing[,8:60]
```

It is important to check for covariates that have virtually no variablility.
```{r}
nzv <- nearZeroVar(features, saveMetrics=TRUE)
nzv
```
It seems that all the variables have some contribution. 

## Models

Its important to try a couple of models to see is the prediction is accurate.

### Classification tree

The basic model by default will be:
```{r}
set.seed(123)
modFit <- train(features$classe ~ ., data = features, method="rpart")
print(modFit, digits=3)
```

```{r}
print(modFit$finalModel, digits=3)
```


```{r}
library(rattle)
fancyRpartPlot(modFit$finalModel)
```

Now that we have the model we are able to test it in the testing set:

```{r}
predictions <- predict(modFit, newdata = features)
print(confusionMatrix(predictions, features$classe), digits=4)
```

The results are 0.4956 accurate. So we should try to improve the model, including in this case pre-processing and cross validation.

```{r}
modFit <- train(features$classe ~ .,  preProcess=c("center", "scale"), trControl=trainControl(method = "cv", number = 4), data = features, method="rpart")
print(modFit, digits=3)
```

```{r}
predictions <- predict(modFit, newdata = features)
print(confusionMatrix(predictions, features$classe), digits=4)
```




##References

Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.
